Exploring Rural Shrink Smart Through Guided Discovery Dashboards

### 2022-07-20 SDSS Journal Paper/Paper Draft.Rmd
comments/tweaks

Some communities continue to thrive as they lose population because they adapt, maintaining quality of life and community services for residents while investing in the future. 
This process, \emph{smart shrinkage}, is important for rural areas who have experienced shrinking populations for decades. As small rural towns do not have access to data scientists or even the ability to easily leverage data collected locally to support decisions, our research team will provide communities with data about services in small town Iowa in order to assist with developing strategies to improve quality of life for their residents amid shrinking populations \cite{scc}. We hope to allow towns to discover their own data and compare to other similar towns, centering decision-making on data in the context of small-town Iowa life. In the process, we will assess our visualizations to determine which strategies for user interface and interactive graphics design are most useful to empower town leaders to make discoveries in publicly available data assembled with a focus on items that impact rural quality of life.

@@

Data collected from \url{data.iowa.gov} were used to create the SCC dashboard. Most of these datasets are collected on a town/city or county level, requiring us to carefully join data accounting for differences in spatial resolution. \url{data.iowa.gov} contains unique information about residents, including local liquor sales, school building locations, town budgets and expenditures, hospital beds, Medicaid reimbursements, and other details that may provide information about local quality of life.

Data collected from ELSI from \url{https://nces.ed.gov} were used to show the distance to any private or public school. The National Center for Education Statistics (NCES) is the primary federal entity for collecting and analyzing data related to education. These data are used for details are a town level, that will provide information about the quality of life.

Data collected from the Index of Relative Rurality (IRR) \url{https://www.ers.usda.gov/topics/rural-economy-population/rural-classifications/} were used in the SCC dashboard to help classify the towns. The Index of Relative Rurality (IRR) is a continuous, threshold-free, and unit-free measure of rurality. It is an alternative to the traditional discrete threshold-based classifications.The IRR ranges between 0 (low level of rurality, i.e., urban) and 1 (most rural). Four steps are involved in its design:
\begin{enumerate}
\item Identifying the dimensions of rurality: population size, density, remoteness, and built-up area.
\item Selecting measureable variables to adequately represent each dimension:
    \begin{itemize}
        \item Size: logarithm of population size
        \item Density: logarithm of population density.
        \item Remoteness: network distance.
        \item Built-up area: urban area (as defined by the US Census Bureau) as a percentage of total land area.
    \end{itemize}
\item Re-scaling the variables onto bounded scales that range from 0 to 1.
\item Selecting a link function: unweighted average of the four re-scaled variable.
\end{enumerate}

@@

Using this data, we created a dashboard which allows communities to explore these data and compare and contrast their local community to other communities of similar size and location. In addition to manual comparisons created by the user, we will use statistical clustering methods to identify groups of towns which employ similar strategies to maintain resident quality of life.

@@

>>>>

Some communities continue to thrive as they lose population because they adapt, maintaining quality of life and community services for residents while investing in the future. 
This process, \emph{smart shrinkage}, is important for rural areas who have experienced shrinking populations for decades. 
As small rural towns do not have access to data scientists or even the ability to easily leverage data collected locally to support decisions, our research team will provide communities with data about services in small town Iowa in order to assist with developing strategies to improve quality of life for their residents amid shrinking populations \citep{scc}. 
We hope to allow towns to discover their own data and compare to other similar towns, centering decision-making on data in the context of small-town Iowa life. 
In the process, we will assess our visualizations to determine which strategies for user interface and interactive graphics design are most useful to empower town leaders to make discoveries in publicly available data assembled with a focus on items that impact rural quality of life.
\svp{This paragraph is pretty good, but see if you can tweak it to blend in a bit better with the edits I made above.}

@@

Data collected from \url{data.iowa.gov} were used to create the SCC dashboard. Most of these data sets are collected at a town/city or county \svp{spatial resolution, requiring us to carefully join data to ensure that these differences are respected while collating relevant information at the city level}. \svp{In addition to the more commonly available statistics derived from e.g. the census and American Community Survey,} \url{data.iowa.gov} contains \svp{several} unique \svp{data sets}, including local liquor sales, school building locations, town budgets and expenditures, hospital beds, Medicaid reimbursements, and other details that may provide information about local quality of life.

\svp{Data available on Iowa's data portal were augmented in some cases with higher-quality data sets in cases where the Iowa data were out of date or insufficiently accurate.}
Data collected from ELSI from \url{https://nces.ed.gov} were used to show the distance to any private or public school. The National Center for Education Statistics (NCES) is the primary federal entity for collecting and analyzing data related to education.
\svp{Cite Dave's papers about the ISTP QOL metrics, including education}
<!-- These data are used for details are a town level, that will provide information about the quality of life. -->

Data collected from the Index of Relative Rurality (IRR) \svp{Create a bibtex entry for this and cite it properly, please. Same goes for most of the other URLs.} \url{https://www.ers.usda.gov/topics/rural-economy-population/rural-classifications/} were used in the SCC dashboard to help classify the towns. The Index of Relative Rurality (IRR) is a continuous, threshold-free, and unit-free measure of rurality. It is an alternative to the traditional discrete threshold-based classifications.The IRR ranges between 0 (low level of rurality, i.e., urban) and 1 (most rural). Four steps are involved in its design:
\begin{enumerate}
\item Identifying the dimensions of rurality: population size, density, remoteness, and built-up area.
\item Selecting measureable variables to adequately represent each dimension:
    \begin{itemize}
        \item Size: logarithm of population size
        \item Density: logarithm of population density.
        \item Remoteness: network distance.
        \item Built-up area: urban area (as defined by the US Census Bureau) as a percentage of total land area.
    \end{itemize}
\item Re-scaling the variables onto bounded scales that range from 0 to 1.
\item Selecting a link function: unweighted average of the four re-scaled variable.
\end{enumerate}

@@

@@

<!-- Not sure this paragraph is in the right place and/or necessary -->
<!-- Using this data, we created a dashboard which allows communities to explore these data and compare and contrast their local community to other communities of similar size and location. In addition to manual comparisons created by the user, we will use statistical clustering methods to identify groups of towns which employ similar strategies to maintain resident quality of life. -->
